---
title: "Documentation for Natura 2000 Data Comparison Script"
author: "Emmanuel Oceguera-Conchas"
date: "2023-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This documentation provides instructiones for using the R script that compares spatial and tabular data fot Natura 2000 sites. 
The script helps identify missing or duplicates sites codes between the two datasets.

## Prerequisites

Before using this script, ensure you have the following:

- R installed on your system
- Required R packages: `stringr`, `sf`, `tidyverse`, and `dplyr`.
- The spatial data file for Natura 2000 sites (ESRI Shapefile format).
- The tabular data file for Natura 2000 sites (CSV format).

## Script usages

1. Open R and set your work directory using `setwd()` function if necesary. Replace `"YOUR_DIRECTORY_PATH"` with the appropiate directory path.

```R

# Set work directory if needed
setwd("YOUR_DIRECTORY_PATH")

```
2. Load the required libraries by running the following commands:

```R

# Load libraries
library(stringr)
library(sf)
library(tidyverse)
library(dplyr

```
3. Define the file paths for your spatial and tabular data:

```R
natura_2000_path <- "path/to/spatial/data.shp"
natura_2000_tab_path <- "path/to/tabular/data.csv"

```

it has to be continued here, just continues explane the script as above.

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
